Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-14 aistats.tex
clean up locally quadratic

\subsection{Asymptote}
\label{ssec:asymptote}
As a reference point for any finite convergence rate, it is interesting to know the asymptotic behavior of these quantities as $n \rightarrow +\infty$.
We now review classical asymptotic results.
We point the reader towards \citet[\S1.1]{ostrovskii2021finite} for a more comprehensive overview.
Asymptotic results typically show that the MLE expressed with natural parameters $\nat$ is asymptotically normal: it reaches the Cramér-Rao lower bound \citep[for instance Ch4.2]{vdv1998asymptotic}.
With exponential families, the MAP is more simply expressed with mean parameters $\meanp$.
We get the asymptote by approximating the Bregman divergence using a second order Taylor expansion
\alignn{
    \bregmanconj(\m^* ; \m)
    &= \frac{\norm{\m^* - \m}^2_{\mF}}{2}
    + O(\norm{\m - \m^*}^3),
    \label{eq:bregmanTaylor}
}
where the Mahalanobis norm  $\| x \|_{\mF}^2 = x^\top \mF x$  is induced by $\mF  := \nabla^2\conj(\m^*)$, the hessian of the entropy at the optimum. It happens that  $\mF$ is also the inverse \textit{Fisher information matrix} at $\nat^*$, since
\aligns{
    \mF
    :=\nabla^2\conj(\m^*)
    = \nabla^2\logpart(\nat^*)^{-1}
    = \Cov_{\nat^*}[T(X)]^{-1}  \; .
}
Plugging~\eqref{eq:defMLE} or~\eqref{eq:defMAP} into~\eqref{eq:bregmanTaylor}, one can show that MLE and MAP verify
\begin{align}
	\label{eq:asymptote}
	\E \bregmanconj \left (\E [T(X)] ; \hat \meanp_n^\text{MLE/MAP} \right )
	= \frac{d}{2n} + O(n^{- \frac{3}{2}}) \; .
\end{align}
Detailed derivation is in~\cref{app:asymptote}
Both MLE and MAP have the same asymptote, as the contribution of the prior $n_0 \meanp_0$ gets negligible compared to the data for large $n$.
This asymptote is also independent of the optimum $\meanp^*$.

\subsection{Quadratic Case}
\label{ssec:quadratic}
As another classical reference point, let us consider the case $\logpart(\nat) = \half \norm{\nat}_2^2$.
For instance, this is the log-partition of a Gaussian with known variance $I$,
\[
	\cX=\real^d,\quad \nu(dx) = \exp\paren{\textstyle \half[-\|x\|^2]} dx,\quad T(x)=x.
\]
Then $\conj(\meanp) = \half \norm{\meanp}_2^2$ as well, and both Bregman divergences are squared $\ell^2$ distances:
\begin{align}
	\bregmanconj(\meanp^* ; \meanp) = \half \norm{\meanp^* -  \meanp }_2^2  \; .
\end{align}
Thanks to the independence of samples, we can break down the MLE into individual point's contributions
\begin{align}
	\expect{\half \norm{\m^* -  \inv{n}  \smallsum_i T_i}_2^2}
	=\frac{\Var(T)}{2n}
	=\frac{1}{2n}
\end{align}
Adding a reference mean $\m_0$ to get the MAP yields to
\begin{align}
	\expect{\bregmanconj(\meanp^*; \MAPm)}
	&= \frac{n \Var(T) +  n_0^2 \norm{\m^* -  \m_0}^2}{2(n+n_0)^2}.
	%\\ &= O\left(\frac{\Var(T)}{n} \right) + O\left(\frac{\norm{\m^* -  \m_0}^2}{n^2} \right)
	\label{eq:MAP_quadratic}
\end{align}
In both cases, we have a variance term in $O(n^{-1})$ and a bias term in $O(n^{-2})$. Ideally, such result should also similarly for arbitrary exponential families.
If we make restrictive assumptions on the log-partition function, we can relate $\bregmanconj$ to a norm or a quadratic.

{\bf If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    &\leq L \norm{\m^* - \m} + \norm{\nat} \norm{\m^* - \m} \\
    &\leq 2L \norm{\m^* - \m}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz, and we get a $O(\inv{\sqrt{n}})$ rate.
Note that none of the standard exponential families verify this assumption.
\TODO{connection with optim non-smooth rate.}

{\bf If $\conj$ is $L$-smooth}\footnote{e.g. $\nabla\conj$ is $L$-Lipschitz.} (e.g. $\logpart$ is $\frac{1}{L}$-strongly convex \citep{kakade2009duality}), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    \leq \frac{L}{2} \norm{\m^* - \m}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic, and we get~\eqref{eq:MAP_quadratic} as an upper bound.
Besides the Gaussian with known variance, no standard exponential family verify this assumption either.
Note that it is also possible to get (more complex) upper bounds under restricted notions of strong-convexity \citep{negahban2012unified}.
\TODO{connection with optim smooth rate.}

\subsection{Locally Quadratic Case}
\label{ssec:local-quadratic}
FromTaylor expansion~~\eqref{eq:bregmanTaylor},
we know that all Bregman divergences are locally quadratic.
Under some assumptions, such as self-concordance of $\conj$\footnote{
In 1d, $f$ is self-concordant iff $\forall x, \abs{f'''(x)} \leq 2 \abs{f''(x)}^{\frac{3}{2}}$.
} \citep[Ch.4.1]{nesterov2003introductory}, we can quantify when this quadratic behavior kicks in. Proofs for this subsection are in \cref{app:self-concordant}.
\begin{proposition}
Let $\conj:\cM\rightarrow \real$ be a self-concordant convex function, $\m, \m^* \in\cM$ and $\mF = \nabla \conj(\m^*)$. Then
\aligns{
	\norm{\m^*-\m}_{\mF} < 0.21
	\implies
	\bregmanconj(\m^*,\m) \leq \norm{\m^*-\m}_{\mF}^2
}
\end{proposition}
\footnote{$0.21$ is a value of $x$ such that $x^2 \geq -\frac{x}{1-x} - \log(1 - \frac{x}{1-x})$}.
To gain insights into how many samples are needed, we can estimate when $\expecti{\norm{\m^*-\m}_{\mF}} < 0.21 $.
For the MLE, $\expecti{\norm{\m^*-\hat \m_n}_{\mF}} = \frac{d}{n}$, so a sufficient condition is $n \geq 5 d$.
For MAP, using~\eqref{eq:MAP_quadratic}, we get the sufficient condition $n\geq 5d + \sqrt{5}\norm{\m^* -  \m_0} - n_0$.
This means that on average, we need much more samples than the dimension to reach the quadratic regime.

We know that $\conj$ is self-concordant when $T$ is 1-dimensional and $\logpart$ is self-concordant.
This is true for
exponential distributions,
Gaussians with known mean,
Laplace with known mean,
Pareto with known minimum value,
or Weibull with known shape $k$.
The entropy is also self-concordant when $T$ lives in a compact \citep{bubeck2015entropic}.
This is true for categorical or Dirichlet distributions.

With some more work, it is possible to get a wealth of results under this assumption.
For instance \citet{ostrovskii2021finite} characterizes the number of samples needed to reach this quadratic behavior for arbitrary parametric models, under the assumptions of self-concordance, or pseudo self-concordance \citep{bach2010self}.
Or in the world of exponential families, \citet{kakade2010learning} assumes a local bound on all higher order moments of $\logpart$ in $\nat^*$, to quantify when $\logpart$ can be sandwiched between quadratics in $\nat$, but this does not directly translate into quadratics in $\m$.

All these works give \textit{large sample} results -- results that hold only for $n\geq N$ for some constant $N$ -- but none of them  apply to small $n$.

\citet{anastasiou2017bounds} and
\citet{marteauferey2019beyond})
\fdk{
We know that if the problem is strongly convex (in $\norm{\cdot}_2$), averaged SGD achieves $O(1/n)$ rate.
[Robust stochastic approximation approach to stochastic programming]. (The paper uses a bounded stochastic gradient assumption, but can be relaxed to bounded variance instead.) However, strong-convexity in the Euclidean norm is not strictly necessary to find an algorithm with convergence rate of $O(1/n)$, as shown by [Non-strongly-convex smooth stochastic approximation with convergence rate O(1/n)] for linear and logistic regression. Their proof techniques based on self-concordance have been extended to other setting, [Finite-sample analysis of M-estimators using self-concordance, Learning Exponential Families in High-Dimensions: Strong Convexity and Sparsity].
The same tools are used in statistics, and evething essentially boils down to controlling how different we are from a Gaussian rv, that we're going to get by the CLT, like [Bounds for the normal approximation of the maximum likelihood estimator, Anastasiou].
}

@@

As introduced in the technical background, MAP can be interpreted as stochastic mirror descent (SMD).
This means that we may obtain a convergence rate for MAP from an optimization analysis, 
and any insights gained from the MAP may inform further design and analysis of SMD.
We first review the assumptions of relative smoothness, useful to deal with non-smooth functions, 
before investigating 3 analysis of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{Relative Smoothness}
Mirror descent (MD) \citep{nemirovski1983problem,beck2003mirror}, also known as
Bregman (proximal) gradient, relative gradient descent or NoLips,
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD \citep{nemirovski2009robust,ghadimi2012optimal}
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2(MD) \& Th.6.3(SMD)]{bubeck2015convex}.
In our case, these assumptions do not hold.
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

Recently, these assumptions have been relaxed to the $\stgcvx$-strong convexity and $\smooth$-smoothness of $f$
\emph{relative} to a reference function $\logpart$, defined as
\aligns{
	\stgcvx \cB_{A}(x, y)
	\leq
	\cB_f(x,y)
	\leq
	\smooth \cB_A(x,y) \; .
}
When $\logpart = \norm{\cdot}^2$, we recover standard (Euclidean) smoothness, and gradient descent.
Those conditions ensure the linear convergence of MD with mirror map $\nabla A$,
extending the theory to non-smooth functions $f$ and non-strongly convex potential $\logpart$
\citep{birnbaum2011distributed, bauschke2017descent, lu2018relatively}.

For exponential families, the log-likelihood perfectly fit into this framework, as
\aligns{
	f(\theta) = A(\theta) - \expect{\lin{T(X), \theta}}
}
is both $1$-smooth and $1$-strongly convex relative to $A$. 
So we are left with finding an applicable convergence rate for SMD under relative assumptions.
As we investigate in the next section, several pieces of work have addressed this question, under various assumptions on the stochasticity \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}.

@@

\textbf{Caveat:} these rates apply to the Bregman in the other direction, e.g. $\bregman(\nat^*, \MAPt)$, and sometimes not on the last iterate, but with tail averaging in the primal.

@@

\subsubsection{Bounded Variance}
\citet{hanzely2018fastest} assume the covariance between negative gradient and next iterate is upper bounded at every time step: $\forall n \geq 1, \forall \hat \nat_{n-1},$
\alignn{
	\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) \leq \lr_n C
	\label{eq:hanzely}
}
for some constant $C$.
For a Gaussian with known variance ($A(\theta) = \frac{1}{2}\norm{\theta}^2$),
this definition recovers the variance of the stochastic gradient
\[
	\expect[\tilde g_t\!\!]{\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2}\leq C.
\]
More generally, if $A$ is $\stgcvx$-strongly convex w.r.t. $\norm{\cdot}$,
the LHS of \cref{eq:hanzely} is bounded by
$\expecti{\frac{1}{\stgcvx}\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2_*}$.
Then a bound on the variance of stochastic gradients is sufficient for \cref{eq:hanzely}
to hold.
%Another perspective on this covariance is as the expectation of the symmetrized Bregman $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$ between stochastic and deterministic updates
%\aligns{
%\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) = \E[\cS(\MAPm - \lr g_n  )]
%}
%thus showing positivity.
Under this assumption, \citet[Lem.4.8]{hanzely2018fastest} prove a $O(1/n)$ convergence rate with $O(1/n)$ step-sizes, 
by applying tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.
But in general, this assumption rarely holds.
For the MAP, this translates into $\forall n \geq 1,$
\alignn{
\expect[T_n]{\lin{T_n - \mu^* , \MAPt}} \leq \lr_n C
}
which fails to hold on the gaussian variance example for instance. \TODO{ DAMIEN short proof or formalize.}

@@

\citet{dragomir2021fast} use the weaker assumption 
\alignn{
	\expect[\tilde g]{
		\cB_{A^*}(\MAPm - 2\lr g(\theta_*), \MAPm)
	} \leq 2 \lr^2 C \; ,
	\quad
	\forall n \geq 1.
	\label{eq:dragomir}
}
In words, this bounds the variance of the gradients at the minimum, but with a metric that depends on the iterate $\MAPm$, and this bound should hold for every iterate we may encounter on the stochastic trajectory.
% As observed by Frederick, this is really weaker than Hanzely, because it is only the first half of their assumption. There is only a factor 2 that's preventing a direct inequality.
Under this assumption, they prove convergence up to a variance ball with a constant step-size.
Their descent lemma \citep[Eq.(12)]{dragomir2021fast} 
is the analog of a modern analysis of SGD~\citep[Th.3.2]{gower2019sgd}, 
meaning we can plug in a $O(1/n)$ decreasing step-size and obtain a $O(1/n)$ convergence rate.

This seems to be a solution to our problem, but it suffers from the same issues as \citet{hanzely2018fastest}. 
For barrier objectives, \cref{eq:dragomir} does not hold uniformly on every path of iterates, 
because these paths can get arbitrarily close to boundaries where $\bregmanconj$ diverges.

>>>>

\subsection{Asymptote}
\label{ssec:asymptote}
As a reference point for any finite convergence rate, it is interesting to know the asymptotic behavior of these quantities as $n \rightarrow +\infty$.
We now detail a classical asymptotic result.
Proofs are in \cref{app:asymptote}, and \citet[\S1.1]{ostrovskii2021finite} offers a great review.

Statistics typically give results on $\nat$, but the MAP~\eqref{eq:defMAP} is more simply expressed with $\meanp$, so let us focus on $\bregmanconj$.
Bregman divergences are locally quadratic, as seen via a second order Taylor expansion
\alignn{
    \bregmanconj(\m^* ; \m)
    &= \frac{\norm{\m^* - \m}^2_{\mF}}{2}
    + O(\norm{\m - \m^*}^3),
    \label{eq:bregmanTaylor}
}
where the Mahalanobis norm  $\| x \|_{\mF}^2 = x^\top \mF x$  is induced by $\mF  := \nabla^2\conj(\m^*)$, the hessian of the entropy at the optimum. It happens that  $\mF$ is also the inverse \textit{Fisher information matrix} at $\nat^*$, since
\aligns{
    \mF
    :=\nabla^2\conj(\m^*)
    = \nabla^2\logpart(\nat^*)^{-1}
    = \Cov_{\nat^*}[T(X)]^{-1}  \; .
}
Plugging the MLE~\eqref{eq:defMLE} or MAP~\eqref{eq:defMAP} into~\eqref{eq:bregmanTaylor}, one can show  that 
\begin{align}
	\label{eq:asymptote}
	\E \bregmanconj \left (\E [T(X)] ; \hat \meanp_n^\text{MLE/MAP} \right )
	= \frac{d}{2n} + O(n^{- \frac{3}{2}}) \; .
\end{align}
Both MLE and MAP have the same asymptote, as the contribution of the prior $n_0 \meanp_0$ gets negligible compared to the data for large $n$.
For well-specified models, this asymptote is also independent of the optimum $\meanp^*$. or $\mF$.
Let us see another example for which it is trivial to get a rate matching~\eqref{eq:asymptote}.

\subsection{Quadratic Case}
\label{ssec:quadratic}
As another classical reference point, let us consider the case $\logpart(\nat) = \half \norm{\nat}_2^2$.
For instance, this is the log-partition of a Gaussian with known variance $I$,
\[
	\cX=\real^d,\quad \nu(dx) = \exp\paren{\textstyle \half[-\|x\|^2]} dx,\quad T(x)=x.
\]
Then $\conj(\meanp) = \half \norm{\meanp}_2^2$ as well, and both Bregman divergences are squared $\ell^2$ distances:
\begin{align}
	\bregmanconj(\meanp^* ; \meanp) = \half \norm{\meanp^* -  \meanp }_2^2  \; .
\end{align}
Thanks to the independence of samples, we can break down the MLE into individual point's contributions
\begin{align}
	\expect{\half \norm{\m^* -  \inv{n}  \smallsum_i T_i}_2^2}
	=\frac{\Var(T)}{2n}
	=\frac{d}{2n}
\end{align}
Adding a reference mean $\m_0$ to get the MAP yields
\begin{align}
	\expect{\bregmanconj(\meanp^*; \MAPm)}
	&= \frac{n \Var(T) +  n_0^2 \norm{\m^* -  \m_0}^2}{2(n+n_0)^2}.
	%\\ &= O\left(\frac{\Var(T)}{n} \right) + O\left(\frac{\norm{\m^* -  \m_0}^2}{n^2} \right)
	\label{eq:MAP_quadratic}
\end{align}
So we see a variance term defining the $\frac{d}{2n}$ asymptote and a bias term in $O(n^{-2})$.
Can we generalize these results to other families ?
If we make restrictive assumptions on the log-partition function, we can relate $\bregmanconj$ to a norm or a quadratic.

{\bf If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    &\leq L \norm{\m^* - \m} + \norm{\nat} \norm{\m^* - \m} \\
    &\leq 2L \norm{\m^* - \m}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz, and~\eqref{eq:MAP_quadratic} yields a $O(\inv{\sqrt{n}})$ rate.
However none of the common exponential families verify this assumption.

{\bf If $\conj$ is $L$-smooth}\footnote{$\conj$ is $L$-smooth iff $\nabla\conj$ is $L$-Lipschitz.} (e.g. $\logpart$ is $\frac{1}{L}$-strongly convex \citep{kakade2009duality}), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    \leq \frac{L}{2} \norm{\m^* - \m}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic, and we get~\eqref{eq:MAP_quadratic} as an upper bound.
It is also possible to get (more complex) upper bounds under restricted notions of strong-convexity \citep{negahban2012unified}.
The problem is that, besides the Gaussian with known variance, no standard exponential family has a strongly convex log-partition function.
So let us focus on a more realistic set of assumptions in the next section.

\subsection{Locally Quadratic Case}
\label{ssec:local-quadratic}
From Taylor expansion~\eqref{eq:bregmanTaylor},
we know that all Bregman divergences are locally quadratic.
Under some assumptions, such as self-concordance\footnote{
In 1d, $f$ is self-concordant iff $\forall x, \abs{f'''(x)} \leq 2 \abs{f''(x)}^{\frac{3}{2}}$.
} of $\conj$ \citep[Ch.4.1]{nesterov2003introductory}, we can quantify when this quadratic behavior kicks in. Proofs for this subsection are in \cref{app:self-concordant}.
\begin{proposition}
Let $\conj:\cM\rightarrow \real$ be a self-concordant convex function, $\m, \m^* \in\cM$ and $\mF = \nabla \conj(\m^*)$. Then\footnote{$0.21$ is a value of $x$ such that $x^2 \geq -\frac{x}{1-x} - \log(1 - \frac{x}{1-x})$}.
\aligns{
	\norm{\m^*-\m}_{\mF} < 0.21
	\implies
	\bregmanconj(\m^*,\m) \leq \norm{\m^*-\m}_{\mF}^2
}
\end{proposition}
To gain insights into how many samples are needed, we can estimate when $\expecti{\norm{\m^*-\m}_{\mF}} < 0.21 $.
For the MLE, $\expecti{\norm{\m^*-\hat \m_n}_{\mF}} = \frac{d}{n}$, so a sufficient condition is $n \geq 5 d$.
For MAP, using~\eqref{eq:MAP_quadratic}, we get the sufficient condition $n\geq 5d + \sqrt{5}\norm{\m^* -  \m_0} - n_0$.
This means that on average, we need $5$ times more samples than the dimension to reach the quadratic regime and ensure an upper-bound like~\eqref{eq:MAP_quadratic}.

The entropy $\conj$ is self-concordant for a lot of common families.
For instance when $T$ is 1-dimensional and $\logpart$ is self-concordant (App.\ref{app:self-concordant}),
as is the case with:
exponential distributions,
Gaussians with known mean,
Laplace with known mean,
Pareto with known minimum value,
or Weibull with known shape $k$.
The entropy is also self-concordant when $T$ lives in a compact \citep{bubeck2015entropic} -- e.g. categorical and Dirichlet distributions.

There are quite a few pieces of work in this direction. 
\citet{ostrovskii2021finite} characterizes the number of samples needed to reach the asymptotic quadratic behavior, for any parametric models with a self-concordant log-likelihood.
\citet{anastasiou2017bounds} obtains a similar flavor of result under other assumptions on the third derivative of $f$.
More closely, in the world of exponential families, \citet{kakade2010learning} assumes a local bound on all higher order moments of $\logpart$ in $\nat^*$, to obtain a similar result.
However these results are expressed with quadratics in $\nat$, not $\m$, and they do not directly translate to convergence rates for the MAP, but they might with some more work.

These results partially answer~\eqref{problem}. 
Partially, because the present proposition and these related works all give \textit{large sample} results,  that hold when $n\geq N$ for some constant $N$.
None of them apply to small $n$.
That is one caveat. Another one is the gaussian example we saw in \S\ref{ssec:gaussian}, which is not self-concordant.
A full solution to~\eqref{problem} would address both of these caveats.
Informed by the properties we have seen so far, let us investigate a bias-variance decomposition to formulate a conjecture.
%If no current approach solves~\eqref{problem}, then let us formulate a conjecture informed by the examples and partial solutions we have s.

@@

As we saw in the technical background, MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform further design and analysis of SMD.
We first review the assumptions of relative smoothness, useful to deal with non-smooth functions, before investigating 3 analysis of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{Relative Smoothness}
Mirror descent (MD) \citep{nemirovski1983problem,beck2003mirror}, also known as
Bregman (proximal) gradient, relative gradient descent or NoLips,
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD
\citep{nemirovski2009robust,ghadimi2012optimal}
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2(MD) \& Th.6.3(SMD)]{bubeck2015convex}.
In our case, none of these assumptions may hold.
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

Recently, these assumptions have been relaxed to the $\stgcvx$-strong convexity and $\smooth$-smoothness of $f$
\emph{relative} to $\logpart$, defined as
\aligns{
	\stgcvx \cB_{A}(x, y)
	\leq
	\cB_f(x,y)
	\leq
	\smooth \cB_A(x,y) \; .
}
When $\logpart = \norm{\cdot}^2$, we recover standard (Euclidean) smoothness, and gradient descent.
Those conditions ensure the linear convergence of MD with $A$ as the reference function,
thus extending the theory to a family of non-smooth functions $f$ and non-strongly convex potential $\logpart$
\citep{birnbaum2011distributed, bauschke2017descent, lu2018relatively}.

For exponential families, the log-likelihood perfectly fit into this framework, as
\aligns{
	f(\theta) = A(\theta) - \expect{\lin{X, \theta}}
}
is both $1$-smooth and $1$-strongly convex relative to $A$. 
So we are left with finding an applicable convergence rate for SMD under relative assumptions.
As we investigate in the next section, several pieces of work have addressed this question, under various assumptions on the stochasticity \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}.

@@

\textbf{Caveat :} all of these rates apply to the Bregman in the other direction, e.g. $\bregman(\nat^*, \MAPt)$, and sometimes not on the last iterate, but with tail averaging in the primal.

@@

\subsubsection{Bounded Variance}
\citet{hanzely2018fastest} assume the covariance between negative gradient and next iterate is upper bounded at every time step: $\forall n \geq 1, \forall \hat \nat_{n-1},$
\alignn{
\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) \leq \lr_n C
\label{eq:hanzely}
}
for some constant $C$.
For a Gaussian with known variance ($A(\theta) = \frac{1}{2}\norm{\theta}^2$),
this definition recovers the variance of the stochastic gradient
\[
	\expect[\tilde g_t\!\!]{\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2}\leq C.
\]
More generally, if $A$ is $\stgcvx$-strongly convex w.r.t. $\norm{\cdot}$,
the LHS of \cref{eq:hanzely} is bounded by
$\expecti{\frac{1}{\stgcvx}\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2_*}$.
Then a bound on the variance of stochastic gradients is sufficient for \cref{eq:hanzely}
to hold.
%Another perspective on this covariance is as the expectation of the symmetrized Bregman $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$ between stochastic and deterministic updates
%\aligns{
%\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) = \E[\cS(\MAPm - \lr g_n  )]
%}
%thus showing positivity.
Under this assumption, \citet[Lem.4.8]{hanzely2018fastest} prove a $O(1/n)$ convergence rate with a $O(1/n)$ step-size, by applying polynomial tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.

In general however, this assumption hardly holds.
For the MAP, this translates into $\forall n \geq 1,$
\alignn{
\expect[T_n]{\lin{T_n - \mu^* , \MAPt}} \leq \lr_n C
}
which fails to hold on the gaussian variance example for instance. \TODO{ DAMIEN short proof or formalize.}

@@

\citet{dragomir2021fast} introduce a much weaker assumption $\forall n \geq 1, \forall \MAPm,$
\alignn{
	\expect[\tilde g]{
		\cB_{A^*}(\MAPm - 2\lr g(\theta_*), \MAPm)
	} \leq 2 \lr^2 C \; .
	\label{eq:dragomir}
}
In words, this bounds the variance of the gradients at the minimum, but with a metric that depends on the iterate $\MAPm$, and this bound should hold for every iterate we may encounter on the stochastic trajectory.
% As observed by Frederick, this is really weaker than Hanzely, because it is only the first half of their assumption. There is only a factor 2 that's preventing a direct inequality.
Under this assumption, they prove convergence up to a variance ball with a constant step-size.
Their descent lemma \citep[Eq.(12)]{dragomir2021fast} is the exact analog  of a modern analysis of SGD \citep[Th.3.2]{gower2019sgd}, meaning we can plug in a $O(1/n)$ decreasing step-size and obtain a $O(1/n)$ convergence rate.

This seems to be a perfect solution to our problem. Yet it suffers from the exact same qualm as \citet{hanzely2018fastest}. For barrier objectives, \cref{eq:dragomir} cannot hold uniformly on every possible path of iterates, because these paths can get arbitrarily close from boundaries where $\bregmanconj$ becomes infinite.
